The seed germination spectrum of alpine plants: a global meta-analysis

### 2020-11-02 doc/manuscript.Rmd
1st major revision 🍃

**Total word count for the main body of the text** 6,343; **Introduction** 1,968; **Materials and Methods** 1,342; **Results** 1,138; **Discussion** 1,896; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0

@@

* Given the micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology, a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on seed germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
* Alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures. Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also to the lowlands. Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
* Globally, overwintering and warm-cued germination are key drivers of germination in alpine habitats. The interplay between germination physiology and seed morphological traits further reflects pressures to avoid frost or drought stress. Our results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.

@@

Alpine environments occur worldwide, above the maximum elevation at which trees can grow naturally [@RN3000; @RN3024; @RN4750]. These treeless habitats are characterized by low temperatures, unstable substrates and short growing seasons [@RN2392]. Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by diaspores (i.e. seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and colonize suitable new sites in response to environmental changes. Plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241]. However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal. For regeneration by seed to be successful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings [@RN4714; @RN2868; @RN4717]. The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691]. In alpine habitats, seed germination will be a crucial test for plant populations coping with climate change, determining whether they will persist or go into extinction.

Natural selection is expected to favour seed germination requirements that reduce the probability of facing environmental conditions which are not appropriate for seedling survival and growth [@RN3375]. Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368]. The propagules of many plant species have also developed seed dormancy, by which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214]. Different degrees of dormancy within a seed population ensure the distribution of seedling emergence across time, a form of bet-hedging offspring survival against unpredictable environments [@RN3065]. Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719]. Given the high micro-climatic heterogeneity of alpine landscapes, and the wealth of local studies on alpine seed ecology (reviewed in the following paragraphs), a current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants. The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats. A major goal for trait-based seed ecology is therefore to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].

@@

Besides patterns of seed dormancy and germination temperature, the response of alpine seeds to diurnal alternating temperatures and light can provide further information on their capacity to detect fine-scale environmental cues. Fluctuations of diurnal temperature decrease with increasing burial depths and the depth profile of this decrease depends on soil type, vegetation cover and snow [@RN1380; @RN4910; @RN4952]. Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909]. Therefore, the response to alternating temperatures and light can indicate seeds whether they are buried deep in the soil, in rock crevices or under snow. The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898]. A study of 445 species from the Qinghai-Tibet plateau found that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376]. This is intriguing given that alpine habitats typically have a strong day-night climatic variation during the growing season. Regarding light and darkness, some local studies have found alpine species to have a preference for germinating in light [@RN4730; @RN4729]. This trait could favour the creation of a persistent soil seed bank [@RN4718; @RN4929] which is advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732]. Indeed, alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945]. Nevertheless, other alpine species seem to prefer dark germination [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008]. Although some germination studies report laboratory experiments conducted with constant and alternating temperatures, or light and dark conditions, they are very much case-dependent, and therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.

@@

We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthesize their conclusions at the global level. All data processing, analysis and manuscript production has been performed in *R* [@RN2315]. The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. (The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review). To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments. The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated. We received data (Table 1) on species from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps. We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).

@@

We obtained seed dormancy class from Baskin & Baskin [-@RN3214] and Rosbakh et al. [-@RN4743]. We simplified the original dormancy classes to five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant. We followed the concept by Vleeshouwers et al. [-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be separated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al. [-@RN2877] which considers morphological and physical dormancy as classes separated from physiological dormancy.

@@

The final dataset used in our meta-analysis contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species. From these, `r MSstrict` were categorized as strict alpine species and `r MSgeneral` as generalist. Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), in both hemispheres. All regions correspond to middle latitudes.

@@

Overall, these results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713]. Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703]. This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899], a mechanism of “seed memory” that facilitates acclimatisation to changing environments [@RN2249], reducing the stratification requirements of seeds produced in a warmer climate.

@@

We also found a general increase of seed germination with higher incubation temperatures. This effect was more pronounced in strict alpine compared to generalist species, confirming the findings of local studies [@RN3266; @RN2371]. It was also more pronounced in small seeded and endospermic species. Small-seeded species have a lower proportion of seedling survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399]. Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these patterns is that alpine plants received selection pressure from damaging spring frosts and evolved a frost avoidance mechanism by germinating at high temperatures [@RN4908; @RN4906; @RN4722]. Moreover, strict alpine species may be less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases (except in Mediterranean and semi-arid mountains), but at the same time there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914]. Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress: small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost, and large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and gives seedlings adequate time to develop a root system to cope with desiccation risks in summer. Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet tolerate heat up to 40–50°C [@RN4902].

In our analysis, increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature. While asynchronous germination is thought to be a form of adaptation to unpredictable environments [@RN3734; @RN3690], our results indicate that asynchronous germination occurs when temperatures are relatively cool, likely as a bet-hedging strategy against the risk of early-spring frost. On the other hand, a fast and synchronised germination at warm conditions (i.e. in late spring) may be another important strategy that increases alpine seedling survival by avoiding drier conditions in summer. The extreme germination uncertainty syndrome, i.e. staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907]. Staggered germination is usually exhibited as a rapid first wave of germination followed by second wave several weeks or months later [@RN2392]. If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].

@@

We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil and certain microsites during the snow-free season [@RN4712; @RN2392; @RN3029]. Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had some effect. This partly contradicts the findings of Liu et al. [-@RN3376], who reported a lack of a response to temperature fluctuation for the species distributed only at high elevations of the Qinghai-Tibet Plateau. Liu and co-workers argued that, in that area, high temperature fluctuations may occur in all months of the year and therefore fluctuating temperature alone may not be a reliable indicator of suitable conditions for seedling establishment and growth, while warm daily temperatures could be a cue that the short summer growing season has begun [@RN3376]. When analysed at the global level, alternating temperatures generally increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season. It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4956; @RN4904].

@@

Using the whole set of traits compiled in this study we visualize the variation in the seed germination spectrum of alpine plants. On one end of the spectrum, strict alpine species tend to show slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy. On the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty) and germination traits that suggest possible fast germination after specific cues (e.g. scarification or alternating temperatures). The dormancy classes without a physiological component (non-dormant, physical, and morphological) appear associated with generalist species. This gradient of regenerative strategies mirrors prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279]. “Fast” regeneration can be understood as a plastic and opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925]. The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome and might be the result of convergent evolution acting on those species that are truly restricted to the alpine vegetation belt.

@@

@@

Alpine plants from different regions of the world respond similarly to germination cues that are expected to support successful regeneration: cold stratification, warm and alternating temperatures, and light. These cues, particularly cold stratification and warm temperatures, are generally associated with cold-adapted herbaceous species in a broad sense [@RN3397; @RN3214; @RN2365]. Nonetheless, our model shows that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt, suggesting an evolutionary convergence of germination patterns in those species truly adapted to the alpine. We can conclude that, while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate germination syndrome.

We were not able to analyse data from tropical alpine regions, highlighting how tropical alpine seeds are underrepresented in the literature. We recommend that tropical alpine regions be the focus of future research efforts, as it is possible that the cold-adapted germination syndrome is not prevalent there. Tropical alpine environments are relatively aseasonal and most plants have been reported as flowering throughout the year [@RN4957]. Rather than being exposed to snow in winter and high temperatures in summer, tropical alpine plants face a year-round stress from night-time frost [@RN2392]. That means that our study is valid for the temperate alpine system, but not necessarily for tropical mountain species pools, which are subjected to different selective pressures.

Another aspect that calls for future efforts is a finer study of germination cues. To produce this meta-analysis, we had to assume some simplifications in order to merge disparate datasets: we considered coarse cues that are routinely recorded in germination tests, like the presence/absence of light, alternating temperatures or cold stratification. New research should aim to decompose these binary cues into finer quantitative variables which provide a more precise environmental signalling, such as the red:far red ratio [@RN604], the length of the photoperiod [@RN1883], the amplitude of the diurnal thermal oscillations [@RN2387] or the length and temperature of cold stratification [@RN2346].

Overall, our first global assessment of the seed germination spectrum in relation with seed morphology supports the existence of trade-offs between sets of regenerative traits [@RN4655], such as the relationship we found between seed/embryo size, germination speed and germination responses to temperature and light. Future studies will need to consider, as we did, evolutionary history as an intrinsic constraint of the seed ecological spectrum.

@@

EFP received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19). BJA was funded by the Spanish Research Agency (PID2019-108636GA/ AEI/ 10.13039/501100011033). EM was supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew. This research was supported by NGSeeds, the Next Generation Seed Ecology, Evolution and Data Science virtual working group.

@@

The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. Upon publication, a version of record of the repository will be deposited in Zenodo. (The repository will be made open when the manuscript is accepted for publication, a zip version has been uploaded to https://drive.google.com/file/d/12Usqi7fOAbOoZ9Uu9wkJt2FANB629LFP/view?usp=sharing for peer-review)

@@

```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of the primary data. The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval. The line of zero effect is shown. When the credible intervals overlap with the zero-effect line, the effect is not significant. Some effects that overlapped with zero (i.e. had no effect) and whose credible intervals were excessively wide are not shown for clarity's sake. The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:seed. A negative interaction (e.g. seed mass and light, in the case of the germination proportion) indicates that, as the interacting factor increases (e.g. seed mass) the response to the germination cue (e.g. light) decreases, in comparison with the main effect.", fig.align = "center", fig.width=7, fig.height=5.5}
cowplot::ggdraw(fig2)
```

```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of the primary data. The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval. The line of zero effect is shown.", fig.align = "center", fig.width=7, fig.height=5}
cowplot::ggdraw(fig3)
```

>>>>

**Total word count for the main body of the text** 6,843; **Introduction** 1,888; **Materials and Methods** 1,517; **Results** 1,132; **Discussion** 2,306; **Number of figures** 5 in colour; **Tables** 1; **Supporting information** 0

@@

* Assumptions about the germination ecology of alpine plants are presently based on individual species and local studies. A current challenge is to synthesize, at the global level, the seed ecological spectrum of alpine plants.
* We meta-analyse primary data from laboratory experiments conducted across four continents (excluding the tropics) and 661 species, to estimate the influence of six environmental cues on germination proportion, mean germination time and germination synchrony; accounting for possible effects of seed morphology (mass, embryo:seed ratio) and phylogeny.
* Most alpine plants show physiological seed dormancy, a strong need for cold stratification, warm-cued germination and positive germination responses to light and alternating temperatures. Species with a distribution limited to the alpine belt have a higher preference for warm temperatures and a stronger response to cold stratification than species whose distribution extends also below the treeline. Seed mass, embryo size and phylogeny have strong constraining effects on germination responses to the environment.
* Globally, overwintering and warm-cued germination are key drivers of germination in alpine habitats. The interplay between germination physiology and seed morphological traits further reflects pressures to avoid frost or drought stress. Our results indicate the convergence, at the global level, of the seed germination spectrum of alpine species.

@@

Alpine environments occur worldwide above the maximum elevation at which trees can grow naturally [@RN3000; @RN3024; @RN4750]. These treeless habitats are characterized by low temperatures, unstable substrates and short growing seasons [@RN2392]. Although most alpine plants can reproduce clonally [@RN2392], sexual reproduction by diaspores (i.e. seeds, fruits, or fruits with extra-ovarian structures attached; hereafter 'seeds') is the main strategy to maintain genetic diversity and to colonize suitable new sites in response to environmental changes. Plant regeneration by seed is a key process that determines whether species will persist in their current habitat or will be able to migrate to new sites [@RN3055]; and it is also fundamental for plant conservation [@RN3763] and seed-based ecological restoration [@RN3507; @RN2241]. However, the short reproductive season of the alpine habitat constrains the phenological timing of flowering, pollination, seed development and seed dispersal. For regeneration by seed to be successful, dispersed propagules must germinate in the most appropriate period to ensure the survival of seedlings [@RN4714; @RN2868; @RN4717]. The physiological process of seed germination is therefore an essential life stage that must be timed to occur when the environment is favourable for subsequent seedling survival and growth [@RN4691]. In alpine habitats, seed germination will also be a crucial test for plant populations coping with global change, determining whether they will persist or go into extinction.

Natural selection is expected to favour seed germination requirements that reduce the probability of facing environmental conditions which are not appropriate for seedling survival and growth [@RN3375]. Thus, germination is mostly regulated by environmental cues related to water availability and soil temperature [@RN3368]. The propagules of many plant species have also developed seed dormancy, by means of which germination is prevented during periods that are only ephemerally favourable, like a short warm spell in the middle of winter [@RN3214]. Different degrees of dormancy within a seed population ensure the distribution of seedling emergence across time, a form of bet-hedging offspring survival against unpredictable environments [@RN3065]. Other germination cues, such as the response to diurnal alternating temperatures [@RN1380] and the need for light or darkness [@RN4656], allow for a fine-scale detection of germination micro-niches and safe sites [@RN4719]. Despite the high micro-climatic heterogeneity of alpine landscapes, current assumptions about germination ecology of alpine plants are mainly based on local studies (reviewed in the following paragraphs). A current challenge is thus to synthesize, at the global level, the seed ecological spectrum of alpine plants. The seed ecological spectrum [@RN4655] is defined as a set of seed traits that determines the ability of plants to disperse, persist, germinate, and establish in different habitats. A major goal for trait-based seed ecology is therefore to study the relationships and trade-offs between different aspects of the seed ecological spectrum [@RN4655], such as the physiological drivers of germination versus key morphological traits like seed mass [@RN4916] and the embryo to seed ratio [@RN3685].

@@

Besides patterns of seed dormancy and germination temperature, the response of alpine seeds to diurnal alternating temperatures and light can provide further information on their capacity to detect fine-scale environmental cues. Fluctuations of diurnal temperature decrease with increasing burial depths and the depth profile of this decrease depends on soil type, vegetation cover and snow [@RN1380; @RN4910; @RN4952]. Solar irradiance in physiologically significant quantities penetrates only the first millimetres of the soil [@RN4909]. Therefore, the response to alternating temperatures and light can indicate seeds whether they are buried deep in the soil, in rock crevices or under snow. The depth of seed burial is crucial for seedling emergence, as small seeds lack the resources to survive until they reach the soil surface [@RN4898]. A study of 445 species from the Qinghai-Tibet plateau found that species from high elevation alpine meadows did not show a significant response to diurnal temperature oscillations [@RN3376]. This is intriguing given that alpine habitats typically have a strong day-night climatic variation during the growing season. Regarding light and darkness, some local studies have found alpine species to have a preference for germinating in light [@RN4730; @RN4729]. This trait is linked with the creation of a persistent soil seed bank [@RN4718; @RN4929] which may be advantageous in the temporally and spatially unpredictable alpine environments [@RN4731; @RN4732]. Indeed, persistent alpine soil seed banks are more frequent than previously supposed, as consistently reported in recent years [@RN4929; @RN4718; @RN4945]. Nevertheless, other alpine species seem to prefer germination in darkness [@RN2943], which may be interpreted as a strategy for detecting safe sites in rock crevices [@RN3008]. Although some germination studies report laboratory experiments conducted with constant and alternating temperatures, or light and dark conditions, they are very much case-dependent, and therefore it is difficult to evaluate whether these factors show a consistent pattern across global alpine ecosystems.

@@

We conducted a meta-analysis of primary data [@RN4734], which consists of pooling together original data from different local studies to synthesize their conclusions at the global level. All data processing, analysis and manuscript production has been performed in *R* [@RN2315]. The original data, plus R code for the analysis and creation of the manuscript using Rmarkdown, can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. To gather data, we contacted research groups which have performed experiments on seed germination of alpine plants, asking them to provide raw data from laboratory germination experiments. The submitted data had to report the studied species, the germination treatments, the number of seeds sown, and the number of seeds germinated. We received data (Table 1) on species from the Cantabrian Mountains of Spain, the European Alps, the Italian Apennines, the Northern Caucasus, the Qinghai-Tibet Plateau of China, the Chilean Andes, and the Australian Alps. We also received data on European alpine species from *Enscobase*, the database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml).

@@

We classified the species in seed dormancy classes following Baskin & Baskin [-@RN3214] and Rosbakh et al. [-@RN4743], using five categories: physiological (germination prevented by the seed’s internal balance of phytohormones), morphological (germination prevented by an embryo which is underdeveloped at the time of dispersal), morphophysiological (both physiological and morphological dormancy present), physical (germination prevented by a water-impermeable seed coat) or non-dormant. We followed the concept by Vleeshouwers et al. [-@RN3254] to define dormancy as an inner seed property (sometimes called "endodormancy"), which should be differentiated from the absence of appropriate germination cues in the environment ("ecodormancy"); and the classification system of Baskin and Baskin [-@RN3261] based on previous work by Nikolaeva et al. [-@RN2877] who consider morphological and physical dormancy as classes separated from physiological dormancy.

@@

The final dataset used in our meta-analysis contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of experimental conditions) contributed by `r MSgroups` research groups, representing `r MSfamilies` seed plant families and `r MSspecies` species. From these, `r MSstrict` were categorized as strict alpine species and `r MSgeneral` as generalist. Seed lots were originally sampled in `r MSregions` alpine regions of the world (Table 1), at middle latitudes of both hemispheres.

@@

Overall, our results confirm that overwintering is essential for promoting seed germination in alpine species, concurring with previous local studies [@RN3393; @RN2868; @RN707; @RN698; @RN3703; @RN2355; @RN3285; @RN2371; @RN4713]. Since global climate change is causing major changes in snow cover duration [@RN4917; @RN4918], it could disrupt natural cold stratification cycles in alpine environments, compromising population viability and indirectly favouring species with no stratification requirement [@RN3703]. This risk will be partly alleviated by the plasticity of physiological seed dormancy in response to the seed maturation temperature [@RN2358; @RN2899], a mechanism of “seed memory” via maternal effects that facilitates acclimatisation to changing environments [@RN2249], reducing the stratification requirements of seeds produced in a warmer climate.

@@

We also found a general increase of seed germination with higher incubation temperatures. This effect was more pronounced in strict alpine compared to generalist species, confirming previous local studies [@RN3266; @RN2371]. It was also more pronounced in small seeded and endospermic species. Small-seeded species have a lower proportion of seedling survival under drought [@RN4530] but may require warmer germination temperatures than large-seeded species, especially in seasonal climates [@RN2399]. Considering that risk reduction is an important selective pressure for the evolution of germination traits [@RN4915], a plausible explanation for these patterns is that alpine plants received selection pressure from damaging spring frosts and evolved a frost avoidance mechanism by germinating at high temperatures [@RN4908; @RN4906; @RN4722]. Moreover, strict alpine species may be less subjected to drought damage during establishment compared to generalist species: as elevation increases, plants face less drought risk because precipitation increases (except in Mediterranean and semi-arid mountains), but at the same time there is a higher probability of early or late season frosts [@RN4897; @RN2392; @RN4914]. Together, these results indicate that germination patterns in alpine species are driven by an interplay of seed size and germination temperature to escape either frost or drought stress: small seeds are being selected for fast germination at warm temperatures, in order to escape unfavourable early-spring frost, and large seeds are being selected for slower germination under cooler conditions, which should maximize emergence during snowmelt (when water availability is highest) and gives seedlings adequate time to develop a root system to cope with desiccation risks in summer. Supporting this view, seedlings of alpine pioneer species have a low frost resistance [@RN4722], yet tolerate heat up to 40–50°C [@RN4902].

In our analysis, increasing temperatures also decreased germination uncertainty, showing that germination synchrony is a plastic trait driven by temperature. While asynchronous germination is thought to be a form of adaptation to unpredictable environments [@RN3734; @RN3690], our results indicate that asynchronous germination occurs when temperatures are relatively cool, likely as a bet-hedging strategy against the risk of early-spring frost. On the other hand, a fast and synchronised germination at warm conditions (i.e. in late spring) may be another important strategy that increases alpine seedling survival by avoiding drier conditions in summer. The extreme germination uncertainty syndrome, i.e. staggering germination so that emergence occurs both before and after winter, is known in alpine species [@RN2392; @RN3285; @RN4907]. Staggered germination is usually exhibited as a rapid first wave of germination followed by a second wave several weeks or months later [@RN2392]. If autumn seedlings grow faster than spring seedlings [@RN4907], this will likely have cascading effects on individual fitness along the life of the plant [@RN3036; @RN2384].

@@

We found a positive germination response to alternating temperatures and light, conditions that mimic the environment in the upper soil and certain microsites during the snow-free season [@RN4712; @RN2392; @RN3029]. Alternating temperatures had a slightly lower effect on strict alpine species, compared with generalist species, but still had some effect. This partly contradicts the findings of Liu et al. [-@RN3376], who reported a lack of a response to temperature fluctuation for the species distributed only at high elevations of the Qinghai-Tibet Plateau. Liu and co-workers argued that, in that area, high temperature fluctuations may occur in all months of the year and therefore fluctuating temperature alone may not be a reliable indicator of suitable conditions for seedling establishment and growth, while warm daily temperatures could be a cue that the short summer growing season has begun [@RN3376]. When analysed at the global level, alternating temperatures generally increase germination of alpine plants, which corresponds well with the fact that most alpine regions have a strong day-night climatic variation during the growing season. It is also not surprising that we detected a negative interaction between seed mass and the germination response to light and alternating temperature regimes, since large seeds can afford to germinate at greater depths (where they are more protected from the alpine environment) and still reach the soil surface [@RN4898; @RN4904].

@@

Using the whole set of traits compiled in this study we can visualize the variation in the seed germination spectrum of alpine plants. At one end of the spectrum, strict alpine species tend to show slow and synchronous germination, require warm temperatures and light to germinate, and need stratification or GA~3~ to break physiological dormancy. At the other end, generalist species tend to have fast germination, showing a bet-hedging strategy (high germination uncertainty) and germination traits that suggest possible fast germination after specific cues (e.g. scarification or alternating temperatures). The dormancy classes without a physiological component (non-dormant, physical, and morphological) appear associated with generalist species. This gradient of regenerative strategies mirrors prominent gradients of fast or slow population dynamics [@RN4922; @RN4921], fast or slow developing leaf traits [@RN4923], or large sets of plant ecological features [@RN2279]. “Fast” regeneration can be understood as a plastic and opportunistic strategy to cope with frequently disturbed habitats such as avalanche ways, steep eroding slopes, regressing glaciers, and riverbeds [@RN4926; @RN4925]. The “slow” regeneration of strict alpine species, on the other hand, seems to be a more specialized syndrome and might be the result of convergent evolution acting on those species that are truly restricted to the alpine vegetation belt.

@@

## Future directions

Our first global assessment of the seed germination spectrum in relation with seed morphology supports the existence of trade-offs between sets of regenerative traits [@RN4655], such as the relationship we found between seed/embryo size, germination speed and germination responses to temperature and light. Future studies will need to consider, as we did, evolutionary history and seed anatomy as intrinsic constraints of the seed ecological spectrum; while the seed germination spectrum we described here will need to be explored at the community level [@RN2352].

This study also highlights the general lack of germination studies on tropical alpine seeds. We were not able to analyse data from tropical alpine regions because they are underrepresented in the literature. Although there is an increasing interest in plant ecology of tropical alpine regions, we could not find seed germination data to be compared with other regions. We thus recommend to focus future research on seed ecology in high-mountain tropical areas, as it is possible that the cold-adapted germination syndrome is not prevalent there. Tropical alpine environments are relatively aseasonal and most plants have been reported as flowering throughout the year [@RN4957]. Rather than being exposed to snow in winter and high temperatures in summer, tropical alpine plants face a year-round stress from night-time frost [@RN2392]. That means that our study is valid for the temperate alpine system, where alpine habitats mainly occur [@RN4750], but not necessarily for tropical alpine systems, which are subjected to different selective pressures.

Another aspect that calls for future efforts is a finer study of germination cues. To produce this meta-analysis, we had to assume some simplifications in order to merge disparate datasets: we considered coarse cues that are routinely recorded in germination tests, like the presence/absence of light, alternating temperatures or cold stratification. New research should aim to decompose these binary cues into finer quantitative variables which provide a more precise environmental signalling, such as the red:far red ratio [@RN604], the length of the photoperiod [@RN1883], the amplitude of the diurnal thermal oscillations [@RN2387] or the length and temperature of cold stratification [@RN2346]. In addition, new field research is necessary to assess germination phenology in response to temperature and water availability on the upper soil layer.

@@

This study demonstrates how alpine plants from different regions of the world respond similarly to germination cues that support successful regeneration: cold stratification, warm and alternating temperatures, and light. These cues, particularly cold stratification and warm temperatures, are generally associated with cold-adapted herbaceous species in a broad sense [@RN3397; @RN3214; @RN2365]. Nonetheless, our results show that the response to stratification and temperature is more prominent in strict alpine species when comparing them with generalist plants that can grow opportunistically in the alpine belt, suggesting an evolutionary convergence of germination patterns in those species truly adapted to alpine ecosystems. We can conclude that, while alpine plants do not show a unique germination syndrome, they do show a more extreme or specialized version of the cold-adapted temperate germination syndrome.

@@

EFP received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19). BJA and EFP were funded by the Spanish Research Agency (PID2019-108636GA/ AEI/ 10.13039/501100011033). EM was supported by the Future Leaders Fellowship - Diversity and Livelihoods, of the Royal Botanic Gardens, Kew. This research was supported by NGSeeds, the Next Generation Seed Ecology, Evolution and Data Science virtual working group.

@@

The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/alpineseeds. Upon publication, a version of record of the repository will be deposited in Zenodo.

@@

```{r fig2, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on germination, according to the MCMC meta-analysis of primary data. The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval. The line of zero effect is shown. When the credible intervals overlap with the zero-effect line, the effect is not significant. Some effects that overlapped with zero (i.e. had no effect) and whose credible intervals were excessively wide are not shown for clarity's sake. The figure shows first the main effect, then the effects for the strict alpine and generalist groups, and then the interaction with seed mass and embryo:seed. A negative interaction (e.g. seed mass and light, in the case of the germination proportion) indicates that, as the interacting factor increases (e.g. seed mass) the response to the germination cue (e.g. light) decreases, in comparison with the main effect.", fig.align = "center", fig.width=7, fig.height=5.5}
cowplot::ggdraw(fig2)
```

```{r fig3, message = FALSE, echo = FALSE, warning = FALSE, fig.pos = "H", fig.cap = "Effect of the random factors on germination, according to the MCMC meta-analysis of primary data. The dots indicate the posterior mean of the effect size, and the brackets its 95 % credible interval. The line of zero effect is shown.", fig.align = "center", fig.width=7, fig.height=5}
cowplot::ggdraw(fig3)
```
